---
title: "MATH 216 Homework 4"
author: "Katherine Hobbs"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Suggested packages; feel free to use more!
library(tidyverse)
library(USAboundaries)
library(maptools)
library(sp)
library(broom)
library(leaflet)
library(dplyr)
library(ggplot2)
library(rgdal)
library(readr)
library(forcats)
```



## Admistrative:

Please indicate

* Who you collaborated with: Amanda Hotvedt
* Roughly how much time you spent on this HW so far: 7
* The URL of the RPubs published URL [here](http://rpubs.com/khobbs/224852).
* What gave you the most trouble:
* Any comments you have:

I feel good about my deliverable and process, apart from the need to hard code the hub coordinates.

## Question 1:

**Question**:  DHL, a German shipping company, has a ton of money to build a
shipping hub in the United States. Where should they build it?

**Deliverable**: A choropleth map of the lower 48 states (no Puerto Rico) where
the color corresponds to the distance of the given area to the shipping hub,
which is marked on the map with a large red dot.

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Perform your computations here:

Social_Explorer_Data <- read_csv("~/Desktop/Senior Year/Data Science/Homeworks/HW-4/R11276882_SL050.csv")
 
Social_Explorer_Data <- Social_Explorer_Data %>%
  rename(county_pop=SE_T001_001)

Social_Explorer_Data <- Social_Explorer_Data%>% 
  mutate(Geo_FIPS_char = as.character(Geo_FIPS))

#Find USA total population

Social_Explorer_Data <- Social_Explorer_Data %>% 
  mutate(USA_pop=sum(county_pop))

USA_total_pop <-306652441
  
#Polygons data
us_counties <- us_counties()
us_counties_tidy <- tidy(us_counties, region="geoid")

#Centroids

centroids <- coordinates(us_counties) %>% 
  tbl_df()

#Meta data 
us_counties_data <- us_counties@data

#Join Centroid and US_Counties Data

Counties_data_2 <- bind_cols(us_counties_data, centroids)

#Join polygons and meta data
us_counties_combined <- left_join(us_counties_tidy, Counties_data_2, by=c("id"="geoid"))

counties_final <- left_join(us_counties_combined, Social_Explorer_Data, by=c("id" = "Geo_FIPS_char"))

#Filter out unwanted states and county
counties_final <- counties_final %>% 
  mutate(county_name = name)

counties_final <- counties_final %>% 
  filter(state_name!= "Alaska") %>% 
  filter(state_name!= "Hawaii") %>% 
  filter(state_name!= "Puerto Rico") %>% 
  filter(state_name!= "Guam") %>% 
  filter(county_name != "Fairfax")

#Make new data set with distinct county rows. Find weighted means for latitute and longitude, weighted by county population. This new data set lost all of the polygon data, so we will go back to using the original data set for the plot itself. Unfortunately, this means we have to take the lat and long values and hard code them.

counties_distinct <- counties_final %>% 
  distinct(county_name, V1, V2, county_pop, group) %>% 
  ungroup() %>% 
  mutate(weighted_lat = county_pop*((V2)/USA_total_pop)) %>% 
  mutate(weighted_long = county_pop*((V1)/USA_total_pop))

#Sum the weighted lat and long values to find a coordinates of the hub

counties_distinct %>% 
  ungroup() %>% 
  summarise(hub_lat = sum(weighted_lat), hub_long = sum(weighted_long))

#Code the lat and long values.

hub_lat <- 40.63
hub_long <- -101.3

counties_final  <- counties_final  %>% 
  mutate(distance= (((lat - hub_lat)^2) + ((long - hub_long)^2))^(1/2))
```


```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Here is a template of a ggplot call; feel free to change this to suit your
# needs. Once your computations above are complete, set the eval=TRUE on the
# line above to tell R Markdown to run this code block.
ggplot(data=counties_final, aes(x=long, y=lat, group=group, fill=distance)) +
  # Plot choropleth polygons, low -> high being colored white -> dark green
  geom_polygon() +
  scale_fill_gradient(low="darkblue", high="white") + 
  # Trace outlines of areas
  geom_path(size=0.1) +
  # Mark hub
  annotate("point", x=hub_long, y=hub_lat, col="red", size=3) +
  # Correct aspect ratio for map
  coord_map(xlim=c(-130, -65),ylim=c(22,50))
```

Based on my analysis, I recommend that DHL build a shipping hub at -101.3 degrees longitude and 40.63 degrees latitutde. This would place the hub in SW Nebraska. For practical purposes, it may be better to place the hub further east in Lincoln or Ohama. 



