---
title: "MATH 216 Homework 4"
author: "Katherine Hobbs"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Suggested packages; feel free to use more!
library(tidyverse)
library(USAboundaries)
library(maptools)
library(sp)
library(broom)
library(leaflet)
library(dplyr)
library(ggplot2)
library(rgdal)
library(readr)
library(forcats)
```



## Admistrative:

Please indicate

* Who you collaborated with: Amanda Hotvedt
* Roughly how much time you spent on this HW so far: 5
* The URL of the RPubs published URL [here](http://rpubs.com/khobbs/224852).
* What gave you the most trouble:
* Any comments you have:

I feel good about my deliverable, but right now the final latitude and longitude values I am getting for my hub are not reasonable as they are too small and not on the map of the US. Therefore, I have put the hub at an arbitrary point for now. Also, we thought we eliminated any repeats in our final data set, but there are still some repeat county inputs we have to address. 

## Question 1:

**Question**:  DHL, a German shipping company, has a ton of money to build a
shipping hub in the United States. Where should they build it?

**Deliverable**: A choropleth map of the lower 48 states (no Puerto Rico) where
the color corresponds to the distance of the given area to the shipping hub,
which is marked on the map with a large red dot.

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Perform your computations here:

Social_Explorer_Data <- read_csv("~/Desktop/Senior Year/Data Science/Homeworks/HW-4/R11276882_SL050.csv")
 
Social_Explorer_Data <- Social_Explorer_Data %>%
  rename(county_pop=SE_T001_001)

Social_Explorer_Data <- Social_Explorer_Data%>% 
  mutate(Geo_FIPS_char = as.character(Geo_FIPS))

#Polygons data
us_counties <- us_counties()
us_counties_tidy <- tidy(us_counties, region="geoid")

#Centroids

centroids <- coordinates(us_counties) %>% 
  tbl_df()

#Meta data 
us_counties_data <- us_counties@data

#Join Centroid and US_Counties Data

Counties_data_2 <- bind_cols(us_counties_data, centroids)

#Join polygons and meta data
us_counties_combined <- left_join(us_counties_tidy, Counties_data_2, by=c("id"="geoid"))

counties_final <- left_join(us_counties_combined, Social_Explorer_Data, by=c("id" = "Geo_FIPS_char"))

#Filter out unwanted states and county
counties_final <- counties_final %>% 
  mutate(county_name = name)

counties_final <- counties_final %>% 
  filter(state_name!= "Alaska") %>% 
  filter(state_name!= "Hawaii") %>% 
  filter(state_name!= "Puerto Rico") %>% 
  filter(state_name!= "Guam") %>% 
  filter(county_name != "Fairfax")

#Select desired variables and eliminate repeat county inputs....unfortunately we realized that there are still some repeat counties!!

FINAL <- counties_final %>% 
  select(county_name, V1, V2, county_pop) %>% 
  distinct(county_name, V1, V2, county_pop)

#Find total USA population according to this data set
FINAL %>% 
  ungroup() %>% 
  summarise(USA_pop = sum(county_pop))

#Total US in this dataset= 202,756,834

USA_pop <- 305355260

#Find weighted means for latitute and longitude, weighted by county population

FINAL <- FINAL %>% 
  mutate(weighted_lat = county_pop*((V2)/USA_pop)) %>% 
  mutate(weighted_long = county_pop*((V1)/USA_pop))

#Find weighted mean lat and long values for the hub

FINAL %>% 
  ungroup() %>% 
  summarise(hub_lat = mean(weighted_lat), hub_long = mean(weighted_long))

#These values don't make sense as long or lat values.
#So for now, I still have an arbitrary central point coded for the sake of showing the map.

hub_lat <- 36

hub_long <- -95

counties_final  <- counties_final  %>% 
  mutate(distance= (((lat - hub_lat)^2) + ((long - hub_long)^2))^(1/2))
```


```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Here is a template of a ggplot call; feel free to change this to suit your
# needs. Once your computations above are complete, set the eval=TRUE on the
# line above to tell R Markdown to run this code block.
ggplot(data=counties_final, aes(x=long, y=lat, group=group, fill=distance)) +
  # Plot choropleth polygons, low -> high being colored white -> dark green
  geom_polygon() +
  scale_fill_gradient(low="darkblue", high="white") + 
  # Trace outlines of areas
  geom_path(size=0.1) +
  # Mark hub
  annotate("point", x=hub_long, y=hub_lat, col="red", size=3) +
  # Correct aspect ratio for map
  coord_map(xlim=c(-130, -65),ylim=c(22,50))
```





